---
title: "Bio Motion - Clean data"
date: "`r Sys.Date()`"
author: "Robin Sifre <robinsifre@gmail.com>"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(readxl)
library(tidyverse)
library(knitr)
library(DT)
library(ggridges)
library(segmented)
library(DiagrammeR)
options(scipen = 2)
```



```{r loaddata, warning=FALSE, message=FALSE, include = FALSE}
datadir = '/Users/sifre002/Box/sifre002/9_ExcelSpreadsheets/05_BioMotion/' # directory with data
wd = '/Users/sifre002/Documents/Code/bio-motion/' # set wd for helper functions

# Eye-tracking data
dat = read_excel( paste(datadir, 'Summary_dat/17-Jul-2020_Balanced_dat.xlsx', sep=''), sheet = "Sheet1") # Balanced ET data 
error_sessions = read_excel( paste(datadir, 'Summary_dat/17-Jul-2020_ErrorLog_FixCounts_BioMotion.xlsx', sep='') ) # .mat files without data

# Demographics data 
asd = read_excel( paste(datadir, 'partic_characteristics/clipLevelInfo_byDxGroup.xlsx', sep=''), sheet = 'ASD')
bap = read_excel( paste(datadir, 'partic_characteristics/clipLevelInfo_byDxGroup.xlsx', sep=''), sheet = 'HR_BAP')
ua = read_excel( paste(datadir, 'partic_characteristics/clipLevelInfo_byDxGroup.xlsx', sep=''), sheet = 'HR_UA')
demo = rbind(asd,bap,ua)

# Gestational age 
ga = read_csv(paste(datadir, 'partic_characteristics/gest_age.csv', sep='') )

# ADOS
ados_hr = read_excel( paste(datadir, 'partic_characteristics/ADOS.xlsx', sep =''), sheet = 'hr_ados')
ados_lr = read_excel( paste(datadir, 'partic_characteristics/ADOS.xlsx', sep =''), sheet = 'lr_ados')

# Mullen
mullen_hr = read_excel( paste(datadir, 'partic_characteristics/mullen.xlsx', sep =''), sheet = 'hr_mullen')
mullen_lr = read_excel( paste(datadir, 'partic_characteristics/mullen.xlsx', sep =''), sheet = 'lr_mullen')

# AVS data
avs = read_csv( paste(datadir, 'AVS/AVS_values.csv', sep='') )

# LR data already cleaned ( session - level )
# lr_dat = read_excel( paste(datadir, 'Low-Risk-SciRep.xlsx', sep=''), sheet = 'CountData')
```


# Merging participant-characteristic data
```{r}
grViz("digraph {
      graph [layout = dot, rankdir = TB]

      node [fontname = Helvetica, shape = rectangle]

      rec1 [label = '1. Import demographcs \n (session #, age, quality, sex, risk, CBE)']
      rec2 [label = '2. Merge with gestational age data \n (Coded in weeks OR categorical)']
      rec3 [label = '3. Merge with 24-month Mullen scores']
      rec4 [label = '4. Merge with 24-month ADOS scores']
      rec1 -> rec2 -> rec3 -> rec4
      
}",
height = 250
      )
```
Missing counts are based on if the individual appeared in a previous step (e.g. 1. demographics), but not a subsequent step (e.g. 2. gestational age data).  

## Notes 
```{r}
# Clean demographic data
demo=demo %>%
  # Clean risk and outcome 
  mutate(Risk=ifelse(Risk=='low', 'LR', Risk),
         Risk=ifelse(Risk=='high', 'HR', Risk),
         CBE=toupper(CBE),
         CBE=ifelse(CBE=='TD'|CBE=='NON-ASD', 'UA', CBE),
         CBE=ifelse(CBE=='BAP/DD', 'BAP', CBE)) %>%
  # Add 0 to matlab_id
  mutate(session2=ifelse( session<10,  paste(0,session,sep=''), as.character(session) ),
         Matlab_session2= paste(`Matlab ID`, session2, sep='_')) %>%
  dplyr::select(-c(session, Matlab_session)) %>%
  dplyr::select(c(Individual,
                  matlabID = `Matlab ID`,
                  Matlab_session = Matlab_session2,
                  Age=age, 
                  Date, Quality, Sex, Risk, CBE,
                  session=session2))
demo = unique(demo) 
```


```{r}
# Clean gestational age data
ga2 = ga %>% 
  # Remove empty rows
  filter(!is.na(Code)) %>% 
  # First try `Gestational Age in Weeks`
  mutate(ga = ifelse(!is.na(`Gestational Age in Weeks`), `Gestational Age in Weeks`, NA)) %>%
  # Then try ind_gest_age, Labor Delivery Before After Due Date Recoded
  mutate(ga = ifelse(is.na(ga) & !is.na(ind_gest_age), ind_gest_age, ga ),
         ga = ifelse(is.na(ga) & !is.na(`Labor Delivery Before After Due Date Recoded`), `Labor Delivery Before After Due Date Recoded`, ga),
         ga = ifelse(is.na(ga) & !is.na(`Labor Delivery Weeks Before Due Date Recoded`), `Labor Delivery Weeks Before Due Date Recoded`, ga)) %>%
  # Create gest age category 
  mutate(ga_cat = ifelse(birth_order ==1 & first_child_less28wks == 'yes', 'less28wks', NA), # extremely preterm
         ga_cat = ifelse(birth_order ==2 & second_child_less28wks == 'yes', 'less28wks', ga_cat),
         ga_cat = ifelse(birth_order ==3 & third_child_less28wks == 'yes', 'less28wks', ga_cat),
         ga_cat = ifelse(birth_order ==4 & fourth_child_less28wks == 'yes', 'less28wks', ga_cat),
         ga_cat = ifelse(birth_order ==5 & fifth_child_less28wks == 'yes', 'less28wks', ga_cat),
         # very preterm
         ga_cat = ifelse(birth_order ==1 & first_child_28_32wks == 'yes', '28_32wks', ga_cat), 
         ga_cat = ifelse(birth_order ==2 & second_child_28_32wks == 'yes', '28_32wks', ga_cat),
         ga_cat = ifelse(birth_order ==3 & third_child_28_32wks == 'yes', '28_32wks', ga_cat),
         ga_cat = ifelse(birth_order ==4 & fourth_child_28_32wks == 'yes', '28_32wks', ga_cat),
         ga_cat = ifelse(birth_order ==5 & fifth_child_28_32wks == 'yes', '28_32wks', ga_cat),
         #  preterm
         ga_cat = ifelse(birth_order ==1 & first_child_32_37wks == 'yes', '32_37wks', ga_cat), 
         ga_cat = ifelse(birth_order ==2 & second_child_32_37wks == 'yes', '32_37wks', ga_cat),
         ga_cat = ifelse(birth_order ==3 & third_child_32_37wks == 'yes', '32_37wks', ga_cat),
         ga_cat = ifelse(birth_order ==4 & fourth_child_32_37wks == 'yes', '32_37wks', ga_cat),
         ga_cat = ifelse(birth_order ==5 & fifth_child_32_37wks == 'yes', '32_37wks', ga_cat), 
         # full term
         ga_cat = ifelse(birth_order ==1 & first_child_more37wks == 'yes', 'more37wks', ga_cat), 
         ga_cat = ifelse(birth_order ==2 & second_child_more37wks == 'yes', 'more37wks', ga_cat),
         ga_cat = ifelse(birth_order ==3 & third_child_more37wks == 'yes', 'more37wks', ga_cat),
         ga_cat = ifelse(birth_order ==4 & fourth_child_more37wks == 'yes', 'more37wks', ga_cat),
         ga_cat = ifelse(birth_order ==5 & fifth_child_more37wks == 'yes', 'more37wks', ga_cat)) %>%
  # Keep relevant cols
  dplyr::select(id=Code, ga, ga_cat)
         
```



Some kids have more than one 24m ADOS. If this is the case, here was how I selected:  
- First choice: ados2-module1.  
- Second choice: ados2-module2.  
- Third choice: ados2-t.  
- Fourth choice: ados-1.  

```{r}
# Clean_ados 
colnames(ados_hr) = c('id', 'ados_measure', 'ados_age', 'ados_verbal', 'ados_SA', 'ados_RRB', 'ados_Total', 'ados_CSS')
colnames(ados_lr) = c('id', 'ados_measure', 'ados_age', 'ados_verbal','ados_SA', 'ados_RRB', 'ados_Total', 'ados_CSS')

ados=rbind(ados_hr %>% mutate(Risk = 'HR'),
      ados_lr %>% mutate(Risk = 'LR'))

# Some kids have ADOSes as multiple time points, and multiple versions of the ADOS at a single timepoint. 
# To filter, 1) Find the time point closest to 24 months, 2) if there are multiple ADOSes at 24m select based on the best option.

ados2 = ados %>%
  # # # 1. Find time point closest to 24m # # #
  # Identify which Age to keep 
  dplyr::select(id, ados_age) %>%
  mutate(months_from_24 = ados_age - 24 ) %>%
  group_by(id) %>%
  summarize(n = n(), 
            closest = min( abs(months_from_24) )) %>% # find ADOS age closest to 24 
  dplyr::select(id, closest) %>%
  ungroup() %>%
  # Merge score_to_keep with ADOS scores
  right_join(ados, ., by = 'id') %>%
  # Re-create months_from_24 for filtering
  mutate(months_from_24 = abs(ados_age - 24)) %>%
  filter(months_from_24 == closest) %>%
  # # # 2. Prioritize ADOS 2 # # # 
  # Rank ADOS versions from first to last choice
  mutate(rank = ifelse(ados_measure == 'ados2-module1', 1, NA),
         rank = ifelse(ados_measure == 'ados2-module2', 2, rank),
         rank = ifelse(ados_measure == 'ados2-t',       3, rank),
         rank = ifelse(ados_measure == 'ados-1',        4, rank),
         rank = ifelse(ados_measure == 'ados-toddler-published', 5, rank)) %>%
  group_by(id) %>%
  dplyr::slice(which.min(rank)) %>%
  dplyr::select(-rank, Risk) %>%
  ungroup() %>%
  # Binarize verbal ‘A1’ —  0/1/2 = verbal ability, 3/4 = nonverbal
  mutate(ados_verbal_bin = ifelse(ados_verbal<=2, 1, NA),
         ados_verbal_bin =ifelse(ados_verbal == 3 | ados_verbal == 4, 0, ados_verbal_bin))
```

The <b>calibrated severity score</b> that should be more standardized and comparable across modules.  
- `r sum(ados2$ados_measure == 'ados-1')` participants were administered the <b>ados-1</b>. Did not create CSS for subdomains for these.  
```{r}
# Function for calculating CSS for the SA and RRB domains of the ADOS-2
# Input: Age, measure, RRB and SA score
source(paste(wd,'helper_functions/Convert_ADOS.R', sep =''))

ados2 = ados2 %>%
  group_by(id) %>%
  # Call function for each id
  summarize(out = list(Convert_ADOS(mod=ados_measure, age_in_months=ados_age, verbal_flag =ados_verbal_bin, SA_raw=ados_SA, RRB_raw=ados_RRB))) %>%
  # Unlist function output and append to data.frame
  mutate(SA_CSS = unlist(lapply(out, '[[', 'SA_CSS')),
         RRB_CSS = unlist(lapply(out, '[[', 'RRB_CSS')),
         conversion_warning = unlist(lapply(out, '[[', 'warning'))) %>%
  ungroup() %>%
  dplyr::select(id, SA_CSS, RRB_CSS, conversion_warning) %>% 
  left_join(., ados2, by = 'id') %>%
  dplyr::select(id, ados_measure, ados_age, ados_verbal, 
                ados_SA, ados_RRB, ados_Total,
                ados_CSS, ados_SA_CSS = SA_CSS, ados_RRB_CSS = RRB_CSS,
                ados_verbal_bin)
```


```{r}
# Clean mullen
mullen = rbind(mullen_hr, mullen_lr )
colnames(mullen) = c('id', 'mullen_type', 'mullen_age', 
                     'gross_motor_t', 'gross_motor_AE',
                     'fine_motor_t',  'fine_motor_AE',
                     'vis_t', 'vis_AE',
                     'rec_lang_t', 'rec_lang_AE',
                     'expr_lang_t', 'expr_lang_AE',
                     'cog_t', 'NVIQ', 'VIQ')

mullen = mullen %>% distinct()
# For kids with more than one score, use the one closest to 24m. 
mullen2 = mullen %>%
  # Identify which Age to keep
  dplyr::select(id, mullen_age) %>%
  mutate(months_from_24 = mullen_age - 24 ) %>%
  group_by(id) %>%
  summarize(n = n(), 
            closest = min( abs(months_from_24) )) %>% # find age closest to 24 
  dplyr::select(id, closest) %>%
  ungroup() %>%
  # Merge score_to_keep with ADOS scores
  right_join(mullen, ., by = 'id') %>%
  # Re-create months_from_24 for filtering
  mutate(months_from_24 = abs(mullen_age - 24)) %>%
  filter(months_from_24 == closest)

setdiff(unique(mullen$id), unique(mullen2$id)) # Make sure that everyone retained a mullen score 
```

# Merging participant-level data with demographics  
```{r}
# # # # # # # # # # # # # # # # # # # # # # # # # # # # 
# Merge participant-characteristics data together
# # # # # # # # # # # # # # # # # # # # # # # # # # # # 

# Create individual- and session-level demographic data structures 
demo_indiv=demo %>% # participant-level variables
  dplyr::select(Individual, id=matlabID, Sex, Risk, CBE) %>%
  unique()
demo_session=demo %>% # session-level variables
  dplyr::select(matlabID=Matlab_session,
                Age, Date, Quality)


```

# Merging participant characteristic data & seeing who is missing data  
## Gestational age 
```{r}
# Merge gestational age 
demo_indiv2 = left_join(demo_indiv, ga2, by =c('Individual'='id'))
missing_ga = demo_indiv2 %>% filter(is.na(ga) & is.na(ga_cat)) %>% dplyr::select(Individual, Risk)

# Make sure they didn't have any data in ga
to_check = ga %>% filter(Code %in% missing_ga$Individual) %>%
  mutate(na_count = apply(., 1, function(x) sum(is.na(x)))) %>%
  mutate(has_data = ifelse(na_count==ncol(ga)-1, 0, 1)) %>%
  filter(has_data == 1) %>%
  dplyr::select(Code)

if (nrow(to_check) > 0) datatable(to_check, caption = 'Ids with gestational age data we may recover')
```
- There were `r nrow(missing_ga)` individuals missing data on gestational age (n=`r sum(missing_ga$Risk=='HR')` HR).  
- Merging resulted in the number of rows changing by: `r abs(nrow(demo_indiv) - nrow(demo_indiv2))`.  

##  Mullen  
List of kids with 2 mullens equidistant from 24 months:  
```{r}
# Who is left with more than one Mullen 
more_than_one_mullen = mullen2 %>%group_by(id) %>%summarize(n=n()) %>%filter(n>1)
mullen2 %>% filter(id %in% more_than_one_mullen$id) %>% dplyr:: select(id, mullen_type, mullen_age)
```
Keep 29-month mullen.  
```{r}
to_drop = which( mullen2$id == '7227-01' & mullen2$mullen_age == 19 )
mullen2 = mullen2[-c(to_drop),]

demo_indiv3 = left_join(demo_indiv2, mullen2, by =c('Individual'='id'))
missing_mullen = demo_indiv3 %>% filter(is.na(mullen_type)) %>% dplyr::select(Individual, Risk)
# Make sure they didn't have any data in mullen
to_check = mullen %>% filter(id %in% missing_mullen$Individual) 

if (nrow(to_check) > 0) datatable(to_check, caption = 'Ids with Mullen data we may recover')
```
- There were `r nrow(missing_mullen)` individuals missing mullen data (n=`r sum(missing_mullen$Risk=='HR')` HR).  
- Merging resulted in the number of rows changing by: `r abs(nrow(demo_indiv2) - nrow(demo_indiv3))`.  

## ADOS
List of kids with 2 ADOS's equidistant from 24 months:  
```{r}
# Who is left with more than one ADOS 
more_than_one_ados = ados2 %>%group_by(id) %>%summarize(n=n()) %>%filter(n>1)
ados2 %>% filter(id %in% more_than_one_ados$id) %>% dplyr:: select(id, ados_age)

demo_indiv4 = left_join(demo_indiv3, ados2, by =c('Individual'='id'))
missing_ados = demo_indiv4 %>% filter(is.na(ados_measure)) %>% dplyr::select(Individual, Risk)
to_check = ados %>% filter(id %in% missing_ados$Individual) 

if (nrow(to_check) > 0) datatable(to_check, caption = 'Ids with ADOS data we may recover')

# Issue with CSS
missing_CSS_subdomain = demo_indiv4 %>% 
  filter(!(Individual %in% missing_ados$Individual)) %>%
  filter(is.na(ados_SA_CSS)) %>% 
  select(Individual, ados_measure)
```
- There were `r nrow(missing_ados)` individuals missing ADOS data (n=`r sum(missing_ados$Risk=='HR')` HR).
- There  `r nrow(missing_CSS_subdomain)` individuals with issues on converting the subdomains.


```{r}
# Select individual-level demo data 
demo_indiv = demo_indiv4 %>% 
  dplyr::select(-c(closest, months_from_24))
```


```{r}
# Clean ET data (run-level)
# create separate id and visit variables
dat = data.frame(dat)
dat$left=ifelse(dat$mov %% 2 == 0, 1, 0)

# Add AVS values
long_ids = seq(424,432,2) # Long AVS movies (the ones analyzed here)
avs2 = avs %>% 
    filter(MovieNum %in% long_ids) %>% 
    dplyr::select(MovieNum, AVS0 = AVS_Diff, clip_order_1, clip_order_2) %>% # AVS0 = Whole movie AVS 
    # Add part 1 AVS 
    left_join(avs, by = c('clip_order_1'='MovieNum')) %>%
    dplyr::select(MovieNum, AVS0, AVS1=AVS_Diff, clip_order_2 = clip_order_2.x) %>%
    # Add part 2 AVS
    left_join(avs, by=c('clip_order_2'='MovieNum')) %>%
    dplyr::select(MovieNum, AVS0, AVS1, AVS2=AVS_Diff) 
# odd movies have the same avs score
avs2=rbind(avs2, avs2 %>% mutate(MovieNum = MovieNum+1)) %>%
  arrange(MovieNum)

# Merge AVS values with dat 
dat=dat%>% left_join(avs2, by=c('mov'='MovieNum'))

```

# Merge all levels of data together
```{r}
# Merge run-level with session
run_dat=left_join(dat, demo_session, by = 'matlabID') %>%
  mutate(id =lapply(strsplit(matlabID, '_'), '[[', 1), # Create id variable for merging w individual-level data 
         id = unlist(id))

source(paste(wd,'helper_functions/correct_age.R', sep =''))

# Calculate corrected age to add to run_dat 
run_dat = run_dat %>% 
  dplyr::select(id, matlabID, Age) %>% 
  distinct() %>%
  # Merge gestational age data 
  left_join(demo_indiv %>% dplyr::select(id,ga, ga_cat), by = 'id' ) %>%
  group_by(matlabID) %>%
  # Call function for correcting age 
  summarize(out = list(correct_age(age_months = Age, gest_age_weeks = ga))) %>%
  mutate(corrected_age = unlist(lapply(out, '[[', 1)) ) %>% 
  ungroup() %>%
  dplyr::select(matlabID, corrected_age) %>% 
  # Merge back with run-level data 
  left_join(., run_dat, by = 'matlabID') 


```

```{r}
write_csv(x = demo_indiv, path = paste(wd, 'analyses/clean_data/individual_chars.csv',sep=''))
write_csv(x = run_dat, path = paste(wd, 'analyses/clean_data/et_dat.csv',sep=''))
```

