---
title: "Bio Motion - Clean data"
date: "`r Sys.Date()`"
author: "Robin Sifre <robinsifre@gmail.com>"
output:
  rmdformats::readthedown:
    highlight: kate
    self_contained: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(readxl)
library(tidyverse)
library(knitr)
library(DT)
library(ggridges)
library(segmented)
library(DiagrammeR)
options(scipen = 2)
```


```{r loaddata, warning=FALSE, message=FALSE, include = FALSE}
datadir = '/Users/sifre002/Box/sifre002/9_ExcelSpreadsheets/05_BioMotion/' # directory with data
wd = '/Users/sifre002/Documents/Code/bio-motion/' # set wd for helper functions

# Eye-tracking data
dat_hr = read_excel( paste(datadir, 'Summary_dat/Balanced_dat_2020-07-09-HRdata.xlsx', sep=''), sheet = "Sheet1") # Balanced ET data 
dat_lr = read_excel( paste(datadir, 'Summary_dat/Balanced_dat_2020-10-26-LRdata.xlsx', sep=''), sheet = "Sheet1") # Balanced ET data 
dat = rbind(dat_hr, dat_lr)
# Error logs from processing
error_sessions_hr = read_excel( paste(datadir, 'Summary_dat/ErrorLog_FixCounts_BioMotion_20200709_HR.xlsx', sep='') ) # .mat files without data
error_sesssions_lr = read_excel( paste(datadir, 'Summary_dat/ErrorLog_FixCounts_BioMotion_20201026_LR.xlsx',sep=''))

# Demographics data 
asd = read_excel( paste(datadir, 'partic_characteristics/clipLevelInfo_byDxGroup.xlsx', sep=''), sheet = 'ASD')
bap = read_excel( paste(datadir, 'partic_characteristics/clipLevelInfo_byDxGroup.xlsx', sep=''), sheet = 'HR_BAP')
ua = read_excel( paste(datadir, 'partic_characteristics/clipLevelInfo_byDxGroup.xlsx', sep=''), sheet = 'HR_UA')
lr = read_excel( paste(datadir, 'partic_characteristics/clipLevelInfo_byDxGroup.xlsx', sep=''), sheet = 'LR')


# Gestational age 
ga = read_csv(paste(datadir, 'partic_characteristics/gest_age.csv', sep='') )

# ADOS
ados_hr = read_excel( paste(datadir, 'partic_characteristics/ADOS.xlsx', sep =''), sheet = 'hr_ados')
ados_lr = read_excel( paste(datadir, 'partic_characteristics/ADOS.xlsx', sep =''), sheet = 'lr_ados')

# Mullen
mullen_hr = read_excel( paste(datadir, 'partic_characteristics/mullen.xlsx', sep =''), sheet = 'hr_mullen')
mullen_lr = read_excel( paste(datadir, 'partic_characteristics/mullen.xlsx', sep =''), sheet = 'lr_mullen')

# AVS data
avs = read_csv( paste(datadir, 'AVS/AVS_values.csv', sep='') )

# LR data already cleaned ( session - level )
# lr_dat = read_excel( paste(datadir, 'Low-Risk-SciRep.xlsx', sep=''), sheet = 'CountData')
```


# About
This script cleans and merges behavioral and eye-tracking data to prep for a multi-level model. 
- <b>Individual-level data</b>: Sex, gestational age, cognitive outcomes (24-month ADOS score, 24-month Mullen score)  
- <b>Visit-level data</b>: Age at eye-tracking visit, session quality, etc.  
- <b>Trial-level data</b>:  Movie number, % looking left side, right side, etc.  


# Data Cleaning
## Demographics
```{r cleandemo, include=TRUE}
# Merge demographics info from the risk groups 
demo = rbind(asd %>%mutate(Sheet='asd'),
             bap%>%mutate(Sheet='bap'),
             ua%>%mutate(Sheet='ua'),
             lr%>%mutate(Sheet='lr'))
# Clean demographic data
demo=demo %>%
  # Clean risk and outcome 
  mutate(Risk=ifelse(Risk=='low', 'LR', Risk),
         Risk=ifelse(Risk=='high', 'HR', Risk),
         CBE=toupper(CBE),
         CBE=ifelse(CBE=='TD'|CBE=='NON-ASD'|CBE=='NON-ASD-DD', 'UA', CBE),
         CBE=ifelse(CBE=='BAP/DD', 'BAP', CBE)) %>%
  # Add 0 to matlab_id
  mutate(session2=ifelse( session<10,  paste(0,session,sep=''), as.character(session) ),
         Matlab_session2= paste(`Matlab ID`, session2, sep='_')) %>%
  dplyr::select(-c(session, Matlab_session)) %>%
  dplyr::select(c(Individual,
                  matlabID = `Matlab ID`,
                  Matlab_session = Matlab_session2,
                  Age=age, 
                  Date, Quality, Sex, Risk, CBE,
                  session=session2,
                  Sheet))
demo = unique(demo) 
```

Make lists of participants to remove for missing demographics
```{r remove_missing_demo}
# Generate lists of IDs w/ missing info 
no_risk = demo %>% 
  filter(is.na(Risk)) %>%
  dplyr::select(matlabID) %>% distinct()
no_CBE = demo %>% 
  filter(is.na(CBE)) %>%
  dplyr::select(matlabID, Risk) %>%
  distinct()
LR_NonTD = demo %>%
  filter(Risk=='LR' & CBE!='UA') %>%
  dplyr::select(matlabID) %>%
  distinct()
missing_risk_or_CBE = unique(c(no_risk$matlabID,no_CBE$matlabID))
LR_NonTD = LR_NonTD$matlabID

demo= demo %>%
  mutate(drop_missing_risk_or_CBE = ifelse(matlabID %in% missing_risk_or_CBE, 1, 0),
         drop_LR_NonTD = ifelse(matlabID %in% LR_NonTD, 1,0))

# Individual and Visit-level demographic info 
demo_clean = demo %>%
  filter(drop_missing_risk_or_CBE == 0,
         drop_LR_NonTD == 0)
demo_indiv=demo_clean %>% # participant-level variables
  dplyr::select(Individual, id=matlabID, Sex, Risk, CBE) %>%
  unique()
demo_session=demo_clean %>% # session-level variables
  dplyr::select(matlabID=Matlab_session,
                Age, Date, Quality)
```

## Gestational age  
Gestational age was recorded in different fields depending on the study.  In some cases, it is continuous (in weeks), in others it is categorical (preterm, very preterm, etc. ). 
```{r clean_gest, echo = TRUE}
# Clean gestational age data
ga_clean = ga %>% 
  # Remove empty rows
  filter(!is.na(Code)) %>% 
  # First try `Gestational Age in Weeks`
  mutate(ga = ifelse(!is.na(`Gestational Age in Weeks`), `Gestational Age in Weeks`, NA)) %>%
  # Then try ind_gest_age, Labor Delivery Before After Due Date Recoded
  mutate(ga = ifelse(is.na(ga) & !is.na(ind_gest_age), ind_gest_age, ga ),
         ga = ifelse(is.na(ga) & !is.na(`Labor Delivery Before After Due Date Recoded`), `Labor Delivery Before After Due Date Recoded`, ga),
         ga = ifelse(is.na(ga) & !is.na(`Labor Delivery Weeks Before Due Date Recoded`), `Labor Delivery Weeks Before Due Date Recoded`, ga)) %>%
  # Create gest age category 
  mutate(ga_cat = ifelse(birth_order ==1 & first_child_less28wks == 'yes', 'less28wks', NA), # extremely preterm
         ga_cat = ifelse(birth_order ==2 & second_child_less28wks == 'yes', 'less28wks', ga_cat),
         ga_cat = ifelse(birth_order ==3 & third_child_less28wks == 'yes', 'less28wks', ga_cat),
         ga_cat = ifelse(birth_order ==4 & fourth_child_less28wks == 'yes', 'less28wks', ga_cat),
         ga_cat = ifelse(birth_order ==5 & fifth_child_less28wks == 'yes', 'less28wks', ga_cat),
         # very preterm
         ga_cat = ifelse(birth_order ==1 & first_child_28_32wks == 'yes', '28_32wks', ga_cat), 
         ga_cat = ifelse(birth_order ==2 & second_child_28_32wks == 'yes', '28_32wks', ga_cat),
         ga_cat = ifelse(birth_order ==3 & third_child_28_32wks == 'yes', '28_32wks', ga_cat),
         ga_cat = ifelse(birth_order ==4 & fourth_child_28_32wks == 'yes', '28_32wks', ga_cat),
         ga_cat = ifelse(birth_order ==5 & fifth_child_28_32wks == 'yes', '28_32wks', ga_cat),
         #  preterm
         ga_cat = ifelse(birth_order ==1 & first_child_32_37wks == 'yes', '32_37wks', ga_cat), 
         ga_cat = ifelse(birth_order ==2 & second_child_32_37wks == 'yes', '32_37wks', ga_cat),
         ga_cat = ifelse(birth_order ==3 & third_child_32_37wks == 'yes', '32_37wks', ga_cat),
         ga_cat = ifelse(birth_order ==4 & fourth_child_32_37wks == 'yes', '32_37wks', ga_cat),
         ga_cat = ifelse(birth_order ==5 & fifth_child_32_37wks == 'yes', '32_37wks', ga_cat), 
         # full term
         ga_cat = ifelse(birth_order ==1 & first_child_more37wks == 'yes', 'more37wks', ga_cat), 
         ga_cat = ifelse(birth_order ==2 & second_child_more37wks == 'yes', 'more37wks', ga_cat),
         ga_cat = ifelse(birth_order ==3 & third_child_more37wks == 'yes', 'more37wks', ga_cat),
         ga_cat = ifelse(birth_order ==4 & fourth_child_more37wks == 'yes', 'more37wks', ga_cat),
         ga_cat = ifelse(birth_order ==5 & fifth_child_more37wks == 'yes', 'more37wks', ga_cat)) %>%
  # Keep relevant cols
  dplyr::select(id=Code, ga, ga_cat)
```

## ADOS  
### 1. Find ADOS collected clostest to 24mo. In cases where visit has more than 1 module, select the best module:   
- First choice: ados2-module1.  
- Second choice: ados2-module2.  
- Third choice: ados2-t.  
- Fourth choice: ados-1.  
```{r clean_ados, echo = TRUE}
# Clean_ados 
colnames(ados_hr) = c('id', 'ados_measure', 'ados_age', 'ados_verbal', 'ados_SA', 'ados_RRB', 'ados_Total', 'ados_CSS')
colnames(ados_lr) = c('id', 'ados_measure', 'ados_age', 'ados_verbal','ados_SA', 'ados_RRB', 'ados_Total', 'ados_CSS')

ados=rbind(ados_hr %>% mutate(Risk = 'HR'),
      ados_lr %>% mutate(Risk = 'LR'))

# Some kids have ADOSes as multiple time points, and multiple versions of the ADOS at a single timepoint. 
# To filter, 1) Find the time point closest to 24 months, 2) if there are multiple ADOSes at 24m select based on the best option.

ados_clean = ados %>%
  # # # 1. Find time point closest to 24m # # #
  # Identify which Age to keep 
  dplyr::select(id, ados_age) %>%
  mutate(months_from_24 = ados_age - 24 ) %>%
  group_by(id) %>%
  summarize(n = n(), 
            closest = min( abs(months_from_24) )) %>% # find ADOS age closest to 24 
  dplyr::select(id, closest) %>%
  ungroup() %>%
  # Merge score_to_keep with ADOS scores
  right_join(ados, ., by = 'id') %>%
  # Re-create months_from_24 for filtering
  mutate(months_from_24 = abs(ados_age - 24)) %>%
  filter(months_from_24 == closest) %>%
  # # # 2. Prioritize ADOS 2 # # # 
  # Rank ADOS versions from first to last choice
  mutate(rank = ifelse(ados_measure == 'ados2-module1', 1, NA),
         rank = ifelse(ados_measure == 'ados2-module2', 2, rank),
         rank = ifelse(ados_measure == 'ados2-t',       3, rank),
         rank = ifelse(ados_measure == 'ados-1',        4, rank),
         rank = ifelse(ados_measure == 'ados-toddler-published', 5, rank)) %>%
  group_by(id) %>%
  dplyr::slice(which.min(rank)) %>%
  dplyr::select(-rank, Risk) %>%
  ungroup() %>%
  # Binarize verbal ‘A1’ —  0/1/2 = verbal ability, 3/4 = nonverbal
  mutate(ados_verbal_bin = ifelse(ados_verbal<=2, 1, NA),
         ados_verbal_bin =ifelse(ados_verbal == 3 | ados_verbal == 4, 0, ados_verbal_bin))
```

### 2. Calculate <b>calibrated severity score</b> that should be more standardized and comparable across modules.  
- `r sum(ados_clean$ados_measure == 'ados-1')` participants were administered the <b>ados-1</b>. Did not create CSS for subdomains for these.  

```{r ados_CSS, echo = TRUE}
# Function for calculating CSS for the SA and RRB domains of the ADOS-2
# Input: Age, measure, RRB and SA score
source(paste(wd,'helper_functions/Convert_ADOS.R', sep =''))

ados_clean = ados_clean %>%
  group_by(id) %>%
  # Call function for each id
  summarize(out = list(Convert_ADOS(mod=ados_measure, age_in_months=ados_age, verbal_flag =ados_verbal_bin, SA_raw=ados_SA, RRB_raw=ados_RRB))) %>%
  # Unlist function output and append to data.frame
  mutate(SA_CSS = unlist(lapply(out, '[[', 'SA_CSS')),
         RRB_CSS = unlist(lapply(out, '[[', 'RRB_CSS')),
         conversion_warning = unlist(lapply(out, '[[', 'warning'))) %>%
  ungroup() %>%
  dplyr::select(id, SA_CSS, RRB_CSS, conversion_warning) %>% 
  left_join(., ados_clean, by = 'id') %>%
  dplyr::select(id, ados_measure, ados_age, ados_verbal, 
                ados_SA, ados_RRB, ados_Total,
                ados_CSS, ados_SA_CSS = SA_CSS, ados_RRB_CSS = RRB_CSS,
                ados_verbal_bin)
```

### 3. Check for kids with 2 ADOS's equidistant from 24 months:  
```{r echo=FALSE}
more_than_one_ados = ados_clean%>%group_by(id) %>%summarize(n=n()) %>%filter(n>1)
ados_clean %>% filter(id %in% more_than_one_ados$id) %>% dplyr:: select(id, ados_age)
```


## Mullen 
```{r clean_mullen, include = TRUE}
# Clean mullen
mullen = rbind(mullen_hr, mullen_lr )
colnames(mullen) = c('id', 'mullen_type', 'mullen_age', 
                     'gross_motor_t', 'gross_motor_AE',
                     'fine_motor_t',  'fine_motor_AE',
                     'vis_t', 'vis_AE',
                     'rec_lang_t', 'rec_lang_AE',
                     'expr_lang_t', 'expr_lang_AE',
                     'cog_t', 'NVIQ', 'VIQ')

mullen = mullen %>% distinct()
# For kids with more than one score, use the one closest to 24m. 
mullen_clean = mullen %>%
  # Identify which Age to keep
  dplyr::select(id, mullen_age) %>%
  mutate(months_from_24 = mullen_age - 24 ) %>%
  group_by(id) %>%
  summarize(n = n(), 
            closest = min( abs(months_from_24) )) %>% # find age closest to 24 
  dplyr::select(id, closest) %>%
  ungroup() %>%
  # Merge score_to_keep with ADOS scores
  right_join(mullen, ., by = 'id') %>%
  # Re-create months_from_24 for filtering
  mutate(months_from_24 = abs(mullen_age - 24)) %>%
  filter(months_from_24 == closest)

setdiff(unique(mullen$id), unique(mullen_clean$id)) # Make sure that everyone retained a mullen score 
```

List of kids with 2 mullens equidistant from 24 months:  
```{r check_mullen, echo=FALSE}
more_than_one_mullen = mullen_clean %>%group_by(id) %>%summarize(n=n()) %>%filter(n>1)
mullen_clean %>% filter(id %in% more_than_one_mullen$id) %>% dplyr:: select(id, mullen_type, mullen_age)
```
Keep 29-month mullen.  
```{r}
to_drop = which( mullen_clean$id == '7227-01' & mullen_clean$mullen_age == 19 )
mullen_clean = mullen_clean[-c(to_drop),]
```




# Merging data
```{r, echo=FALSE}
grViz("digraph {
      graph [layout = dot, rankdir = TB]

      node [fontname = Helvetica, shape = rectangle]

      rec1 [label = '1. Import demographcs \n (session #, age, quality, sex, risk, CBE) \n Remove participant w/ missing Risk and/or CBE']
      rec2 [label = '2. Merge with gestational age data \n (Coded in weeks OR categorical)']
      rec3 [label = '3. Merge with 24-month Mullen scores']
      rec4 [label = '4. Merge with 24-month ADOS scores']
      rec1 -> rec2 -> rec3 -> rec4
      
}",
height = 250
      )
```

## Merge individual-level demographics with gestational age.
```{r merge_gest_age}
# Merge gestational age 
demo_indiv_ga = left_join(demo_indiv, ga_clean, by =c('Individual'='id'))
```
- Merging resulted in the number of rows changing by: `r abs(nrow(demo_indiv_ga) - nrow(demo_indiv))`.  
- Total # rows in `demo_indiv_ga`: `r nrow(demo_indiv_ga)`

## Merge individual-level demo with Mullen  
```{r merge_mullen}
demo_indiv_ga_mullen = left_join(demo_indiv_ga, mullen_clean, by =c('Individual'='id'))
```
- Merging resulted in the number of rows changing by: `r abs(nrow(demo_indiv_ga) - nrow(demo_indiv_ga_mullen))`.  
- Total # rows in `demo_indiv_ga_mullen`: `r nrow(demo_indiv_ga_mullen)`

## Merge individual-level demo with ADOS
```{r merge_ADOS}
demo_indiv_ga_mullen_ados = left_join(demo_indiv_ga_mullen, ados_clean, by =c('Individual'='id'))
```
- Merging resulted in the number of rows changing by: `r abs(nrow(demo_indiv_ga_mullen) - nrow(demo_indiv_ga_mullen_ados))`.  
- Total # rows in `demo_indiv_ga_mullen_ados`: `r nrow(demo_indiv_ga_mullen_ados)`

```{r final_demo_indiv}
demo_indiv = demo_indiv_ga_mullen_ados %>% 
  dplyr::select(-c(closest, months_from_24))
```

# Log of filtered/missing data
## 1. Demographics  
Participants w/ missing Risk Status
```{r}
no_risk$matlabID
```

Participants w/ missing CBE:
```{r}
setdiff(no_CBE$matlabID, no_risk$matlabID) # Don't double-count if missing risk and CBE 
```

LR participants to drop because CBE is not TD 
```{r}
LR_NonTD
```

```{r}
temp1=demo %>% filter(drop_missing_risk_or_CBE == 1) %>% count(matlabID) # each row = a session
temp2=demo %>% filter(drop_LR_NonTD == 1) %>% count(matlabID)
```
- `r nrow(temp1)` participants (`r sum(temp1$n)` visits) removed for missing Risk or CBE.  
- `r nrow(temp2)` participants (`r sum(temp2$n)` visits) removed for LR participant with non-TD CBE.    















## 2. Gestational age
```{r missing_ga}
missing_ga = demo_indiv_ga %>% filter(is.na(ga) & is.na(ga_cat)) %>% dplyr::select(Individual, Risk)

# Make sure they didn't have any data in raw ga
to_check = ga %>% filter(Code %in% missing_ga$Individual) %>%
  mutate(na_count = apply(., 1, function(x) sum(is.na(x)))) %>%
  mutate(has_data = ifelse(na_count==ncol(ga)-1, 0, 1)) %>%
  filter(has_data == 1) %>%
  dplyr::select(Code)

if (nrow(to_check) > 0) datatable(to_check, caption = 'Ids with gestational age data we may recover')
```
- There were `r nrow(missing_ga)` individuals missing data on gestational age (n=`r sum(missing_ga$Risk=='HR')` HR).  
- Keep in analysis for now. 

## 3. Mullen
```{r missing_mullen}
missing_mullen = demo_indiv_ga_mullen %>% filter(is.na(mullen_type)) %>% dplyr::select(Individual, Risk)
# Make sure they didn't have any data in mullen
to_check = mullen %>% filter(id %in% missing_mullen$Individual) 

if (nrow(to_check) > 0) datatable(to_check, caption = 'Ids with Mullen data we may recover')
```
- There were `r nrow(missing_mullen)` individuals missing mullen data (n=`r sum(missing_mullen$Risk=='HR')` HR).  
- Keep them in dataset for now. 

## 4. ADOS
```{r adoscheck_1}
missing_ados = demo_indiv_ga_mullen_ados %>% filter(is.na(ados_measure)) %>% dplyr::select(Individual, Risk)
to_check = ados %>% filter(id %in% missing_ados$Individual) 

if (nrow(to_check) > 0) datatable(to_check, caption = 'Ids with ADOS data we may recover')
```
- There were `r nrow(missing_ados)` individuals missing ADOS data (n=`r sum(missing_ados$Risk=='HR')` HR).

```{r adoscheck_2}
# Issue with CSS
missing_CSS_subdomain = demo_indiv_ga_mullen_ados %>% 
  filter(!(Individual %in% missing_ados$Individual)) %>%
  filter(is.na(ados_SA_CSS)) %>% 
  dplyr::select(Individual, ados_measure)
```
- There  `r nrow(missing_CSS_subdomain)` individuals with issues on converting the subdomains.


# Eye-tracking data 
## Cleaning
```{r clean_et, echo = TRUE}
# Clean ET data (run-level)
# create separate id and visit variables
dat = data.frame(dat)
dat$left=ifelse(dat$mov %% 2 == 0, 1, 0)

# Add AVS values
long_ids = seq(424,432,2) # Long AVS movies (the ones analyzed here)
avs2 = avs %>% 
    filter(MovieNum %in% long_ids) %>% 
    dplyr::select(MovieNum, AVS0 = AVS_Diff, clip_order_1, clip_order_2) %>% # AVS0 = Whole movie AVS 
    # Add part 1 AVS 
    left_join(avs, by = c('clip_order_1'='MovieNum')) %>%
    dplyr::select(MovieNum, AVS0, AVS1=AVS_Diff, clip_order_2 = clip_order_2.x) %>%
    # Add part 2 AVS
    left_join(avs, by=c('clip_order_2'='MovieNum')) %>%
    dplyr::select(MovieNum, AVS0, AVS1, AVS2=AVS_Diff) 
# odd movies have the same avs score
avs2=rbind(avs2, avs2 %>% mutate(MovieNum = MovieNum+1)) %>%
  arrange(MovieNum)

# Merge AVS values with dat 
dat=dat%>% left_join(avs2, by=c('mov'='MovieNum'))
```
- There are `r nrow(dat)` rows of eye-tracking data.  

# Merge all levels of data together  
## Merge eye-tracking data with visit-level data 
```{r merge_et_vislevel}
# Merge run-level with session
run_dat=left_join(dat, demo_session, by = 'matlabID') %>%
  mutate(id =lapply(strsplit(matlabID, '_'), '[[', 1), # Create id variable for merging w individual-level data 
         id = unlist(id))
```

## Merge eye-tracking data with Individual-level data 
```{r}
source(paste(wd,'helper_functions/correct_age.R', sep =''))

# Calculate corrected age to add to run_dat 
run_dat = run_dat %>% 
  dplyr::select(id, matlabID, Age) %>% 
  distinct() %>%
  # Merge gestational age data 
  left_join(demo_indiv %>% dplyr::select(id,ga, ga_cat), by = 'id' ) %>%
  group_by(matlabID) %>%
  # Call function for correcting age 
  summarize(out = list(correct_age(age_months = Age, gest_age_weeks = ga))) %>%
  mutate(corrected_age = unlist(lapply(out, '[[', 1)) ) %>% 
  ungroup() %>%
  dplyr::select(matlabID, corrected_age) %>% 
  # Merge back with run-level data 
  left_join(., run_dat, by = 'matlabID') 
```

```{r}
write_csv(x = demo_indiv, path = paste(wd, 'analyses/clean_data/', Sys.Date(), 'individual_chars.csv',sep=''))
write_csv(x = run_dat, path = paste(wd, 'analyses/clean_data/',Sys.Date(), 'et_dat.csv',sep=''))
```

